---
title: "p8105_hw6_lt2949"
author: "Li Tian"
date: "2023-11-29"
output: github_document
---
```{r}
library(tidyverse)
library(patchwork)
library(modelr)
library(mgcv)

```


## Problem1
```{r}
library(tidyverse)
library(modelr)
```
In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Problem2

```{r}
#Download the data
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |> 
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |> 
  select(name, id, everything())

```

After loading the data, we clean it up 

```{r df_clean}

# selects relevant columns and renames Central Park to cp_ny for simplicity
strapped_weather_df <- weather_df %>% 
  select(name, tmax, tmin) %>% 
  mutate(
    name = str_replace(name, "CentralPark_NY", "cp_ny")
  ) %>% 
  bootstrap(n = 5000) %>%                                 
  mutate(
    model = map(strap, ~lm(tmax ~ tmin, data = .x)),     
    result = map(model, broom::tidy),                    
    stat = map(model, broom::glance)                    
  )

# a cleaned and filtered df containing variables of interest
clean_weather_strap <- strapped_weather_df %>% 
  select(-model, -strap) %>%                              # removed the original strap sample and model
  rename("strap_run" = .id) %>%                           
  unnest() %>%                                           
  select(strap_run, term, estimate, adj.r.squared) %>%   
  mutate(
    term = case_when(term == "(Intercept)" ~ "beta0",     # renamed intercept to beta0
                     term == "tmin" ~ "beta1",            # renamed slope to beta1
                     TRUE ~ as.character(term))
  ) %>% 
  pivot_wider(names_from = term,                         
              values_from = estimate) %>% 
  mutate(
    estimate_log = log(beta0 * beta1)                    
  ) %>% 
  janitor::clean_names()

```

After cleaning and bootstrapping, the `clean_weather_strap` dataset with `r nrow(clean_weather_strap)` records includes `adj_r_squared`, `beta0`, `beta1`, and `estimate_log` for each bootstrap iteration (`strap_run`). We will examine their distributions using density plots.

```{r density_plots}

# density plot of r-squared with 2.5% and 97.5% quantile labeled
r_squared_plot <- clean_weather_strap %>% 
  ggplot(aes(x = adj_r_squared)) +
  geom_density(color = "green", fill = "green", alpha = 0.3) +
  geom_vline(xintercept = c(quantile(pull(clean_weather_strap, adj_r_squared), probs = 0.025),
                            quantile(pull(clean_weather_strap, adj_r_squared), probs = 0.975)),
             linetype = "dashed",
             color = "red") +
  labs(x = "Adjusted R-squared",
       y = "Frequency",
       caption = "Distribution of adjusted R-squared (left) and ln(β̂1 * β̂) (right)")

# density plot of ln(beta0 * beta1) with 2.5% and 97.5% quantile labeled
estimate_log_plot <- clean_weather_strap %>% 
  ggplot(aes(x = estimate_log)) +
  geom_density(color = "lightblue", fill = "lightblue", alpha = 0.3) +
  geom_vline(xintercept = c(quantile(pull(clean_weather_strap, estimate_log), probs = 0.025),
                            quantile(pull(clean_weather_strap, estimate_log), probs = 0.975)),
             linetype = "dashed",
             color = "red") +
  labs(x = "ln(β̂1 * β̂0)",
       y = "Frequency")

# use patchwork to print them side-by-side
r_squared_plot + estimate_log_plot

```

The plot shows R-squared and log(beta0 * beta1) as nearly normal distributions, with a slight left skew in R-squared. The red dashed lines denote the 2.5% and 97.5% quantiles from the `quantile()` function.

```{r quantiles}

# r-sq 95% CI
quantile(pull(clean_weather_strap, adj_r_squared), probs = c(0.025, 0.975))

# est_log 95% CI
quantile(pull(clean_weather_strap, estimate_log), probs = c(0.025, 0.975))

```

As such, we have a 95% CI of [`r quantile(pull(clean_weather_strap, adj_r_squared), probs = 0.025)`, `r quantile(pull(clean_weather_strap, adj_r_squared), probs = 0.975)`] for adjusted R-squared and [`r quantile(pull(clean_weather_strap, estimate_log), probs = 0.025)` , `r quantile(pull(clean_weather_strap, estimate_log), probs  = 0.975)`] for ln(β̂1 * β̂0).

## Problem3 

```{r}
birthweight_df = 
  read_csv("birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) 

birthweight_df

purrr::map(birthweight_df, ~ sum(is.na(.)))

```
This dataset contains `r nrow(birthweight_df)` rows and `r ncol(birthweight_df)` columns, there is no missing data.\

According to a Wikipedia page[https://en.wikipedia.org/wiki/Low_birth_weight] on low birth weight, certain maternal risk factors like younger age, higher parity, and a history of low birth weight (LBW) infants can influence a baby's birth weight. Based on this information, I incorporated mother's age, number of previous live births, and the count of past LBW infants into the model.

The duration of gestation and the presence of malformations are also intuitively connected to birthweight, hence their inclusion in the model.

Additionally, I wanted to explore the impact of sociodemographic factors on birthweight, leading to the inclusion of the mother's race and family income as variables.

```{r}
model_fit_1 = lm(bwt ~ gaweeks + momage + mrace + malform + parity + fincome + pnumlbw, data = birthweight_df)

birthweight_df %>% 
  modelr::add_residuals(model_fit_1) %>%
  modelr::add_predictions(model_fit_1) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(se = F, color = "red", method = "lm")
  labs(
    title = "Predicted vs. Residuals",
    x = "Predicted",
    y = "Residuals"
    ) +
  theme(plot.title = element_text(hjust = 0.5))

# Fit models
model_2 = lm(bwt ~ gaweeks + blength, data = birthweight_df)
model_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_df)
```


```{r}
# Cross validation
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df = 
  cv_df %>% 
  mutate(
   model_fit_1 = map(.x = train, ~lm(bwt ~ gaweeks + momage + mrace + malform + parity + fincome + pnumlbw, data = .x)),
   model_2 = map(.x = train, ~lm(bwt ~ gaweeks + blength, data = .x)),
   model_3 = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))
  ) %>% 
  mutate(
    rmse_model1 = map2_dbl(.x = model_fit_1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = model_2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(.x = model_3, .y = test, ~rmse(model = .x, data = .y))
  )

# Violin plots 
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

The RMSE plot reveals that out of the three models tested, Model 3 performs the best, evidenced by its lowest RMSE value. Contrary to expectations, Model 1, which was informed by information from Wikipedia, did not align well with the data, suggesting that Wikipedia may not be a reliable source for this purpose.

Model 3 encompasses variables such as head circumference, length, and sex, along with all their interaction terms, including the three-way interaction among them.